---
title: "25-07-01-Between-t-test"
author: "Annika Richter"
date: "2025-07-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Within subject linear mixed model
Monte Carlo Simulation comparing for within-subject analysis

Within Subject design taking into account repeated measurements as well as dyad effects minimum sample size for 90% power for an effect size of 0.25:

3 prompts: 130 ppt
4 prompts: 100 ppt
5 prompts: 80 ppt
6 prompts: 70 ppt


Next possible steps:
- Take into account prompt effect
- Take into account asymetric dyad effect (not there in human-Ai condition)
- Test for between participant analysis
```{r}
library(tidyverse)
library(foreach)
library(lmerTest)
```

```{r}
estimate_power_lmm <- function(n_participants,
                               effect_size = 0.25,
                               n_trials_per_condition = 3,
                               alpha = 0.05,
                               n_simulations = 10000,
                               noise_sd = 1) {
  
  if (n_participants %% 2 != 0) stop("n_participants must be even for dyads")

  sims <- foreach(i = 1:n_simulations, .combine = c) %do% {
    
    # Create participant IDs
    participant_ids <- 1:n_participants
    dyad_ids <- rep(1:(n_participants /2), each = 2)
    
    # Simulate full dataset
    df <- expand.grid(
      participant = participant_ids,
      condition = c("A", "B"),
      trial = 1:n_trials_per_condition
    ) %>%
      mutate(
        dyad = dyad_ids[participant],
        condition_numeric = ifelse(condition == "B", 1, 0),
        re_participant = rnorm(n_participants, 0, 0.5)[participant],
        re_dyad = rnorm(n_participants / 2, 0, 0.3)[dyad],
        error = rnorm(n(), 0, noise_sd),
        outcome = 3 +
          effect_size * condition_numeric +
          re_participant +
          re_dyad +
          error
      )
    
    # Fit LMM: random intercept per participant
    model <- tryCatch({
      lmer(outcome ~ condition_numeric + 
             (1 + condition_numeric| participant) +
             (1 | dyad), 
           data = df)
    }, error = function(e) return(NULL))
    
    if (is.null(model) || isSingular(model)) {
      return(NA)
    }
    
    # Extract p-value
    coefs <- summary(model)$coefficients
    if (!"condition_numeric" %in% rownames(coefs)) return(NA)
    
    pval <- coefs["condition_numeric", "Pr(>|t|)"]
    return(as.integer(pval < alpha))
  }
  
  power <- mean(sims, na.rm = TRUE)
  return(power)
}
```

```{r}
estimate_sample_size_lmm <- function(target_power = 0.9,
                                     effect_size = 0.25,
                                     n_trials_per_condition = 3,
                                     alpha = 0.05,
                                     max_n = 200,
                                     n_simulations = 10000,
                                     noise_sd = 1,
                                     verbose = TRUE) {
  
  results <- tibble()
  
  #for (n in seq(150, max_n, by = 10)) {
  for (n in seq(10, max_n, by = 10)) {
    if (verbose) cat("Testing n =", n, "...\n")
    
    power <- estimate_power_lmm(
      n_participants = n,
      effect_size = effect_size,
      n_trials_per_condition = n_trials_per_condition,
      alpha = alpha,
      n_simulations = n_simulations,
      noise_sd = noise_sd
    )
    
    results <- bind_rows(results, tibble(n = n, power = power))
    
    if (power >= target_power) {
      if (verbose) cat("Minimum sample size for", target_power * 100, "% power:", n, "\n")
      break
    }
  }
  
  return(results)
}

```
```{r}
# Estimate power curve
power_curve <- estimate_sample_size_lmm(effect_size = 0.25,
                                        n_trials_per_condition = 3,
                                        max_n = 160)

```




