---
title: "OES_Preprocessing"
author: "Annika Richter"
date: "2026-05-13"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Pre-Processing for Confirmatory Analysis

# 1. Preprocessing:
- Connect to MongoDB using connection string and mongolite package
- Extract Data from MongoDB

From Database Extract 
- Interleaf Measurements: Flow, Social Presence, Panas, Likeability, Flow Feedback, AI_Detection_Rating 
- Participant_ID
- Shared prompt order
- Condition order

Filter out incomplete data:
- Keep only participants that have a complete dataset - e.g. have completed all the chats and have values in all interleaf pages

```{r}
# Load libaries
library(mongolite)
library(tidyverse)
library(lubridate)
library(tidyr)
library(dplyr)
```

```{r}
# Connect to MongoDB
# Set connection string:
connection_string <- "connection_string_to_mongo_db"

# Connect to sessions data
sessions_chatting_castellum <- mongo(collection = "good_sessions_chatting_castellum", db = "llm-flow-experience-backend", url = connection_string)
```

```{r}
#####################################
# Extract Interleaf Measurements
#####################################
# Create an interator
it <- sessions_chatting_castellum$iterate()

# Initialize empty lists for values
interleaf_flow_list <- list()
shared_prompt_list <- list()
shared_condition_list <- list()
participant_id_list <- c()
dyad_id_list <- c()
partner_id_list <- c()

# Loop through all database documents
while (!is.null(doc <- it$one())) {
  # Extract interleaf_flow, prompt and condition list
  # Interleaf_flow as name is here a little bit misleading, as it actually includes a nested list of all interleaf measures, not only flow
  interleaf_flow_list <- append(interleaf_flow_list, list(doc$task$interleaf_flow))
  shared_prompt_list <- append(shared_prompt_list, list(doc$task$dyad$shared_prompt_order))
  shared_condition_list <- append(shared_condition_list, list(doc$task$dyad$shared_condition_order))
  
  #Extract participant_id
  participant_id_list <- c(participant_id_list, doc$participant_id)
  dyad_id_list <- c(dyad_id_list, doc$task$dyad$dyad_id)
  partner_id_list <- c(partner_id_list, doc$task$dyad$partner_id)
}
```
```{r}
all_docs <- sessions_chatting_castellum$find('{}')
```

```{r}
#####################################
# Store into Dataframe
#####################################
df_prompt <- data.frame(
  participant_ID = participant_id_list,
  shared_prompt_order = I(shared_prompt_list)
)
df_condition <- data.frame(
  participant_ID = participant_id_list,
  shared_condition_order = I(shared_condition_list),
  dyad_ID = dyad_id_list,
  partner_ID = partner_id_list
)

df_interleaf_flow <- data.frame(
  participant_ID = participant_id_list,
  interleaf_flow = I(interleaf_flow_list)
)

head(df_prompt)
head(df_condition)
head(df_interleaf_flow)
```

```{r}
df_condition_long <- df_condition %>%
  unnest_longer(shared_condition_order) %>%    # expand list to rows
  group_by(participant_ID) %>%                 # group by participant
  mutate(chat_number = row_number()) %>%       # create chat number per participant
  ungroup()     # each element of the list becomes a row

df_prompt_long <- df_prompt %>%
  unnest_longer(shared_prompt_order) %>%
  group_by(participant_ID) %>%
  mutate(chat_number = row_number()) %>%
  ungroup()

# Preview
df_condition_long
df_prompt_long
```
```{r}
# Merge prompt and condition dataframes
df_conditions <- df_prompt_long %>%
  left_join(df_condition_long, 
            by = c("participant_ID", "chat_number"))

# Preview
head(df_conditions)
```
```{r}
df_interleaf_long <- df_interleaf_flow %>%
  # make one row per chat
  unnest_longer(interleaf_flow) %>%
  
  # spread the elements of each chat (flow, panas, etc.) into columns
  unnest_wider(interleaf_flow) %>%
  
  # add chat number per participant
  group_by(participant_ID) %>%
  mutate(chat_number = row_number()) %>%
  ungroup()

head(df_interleaf_long)
```
```{r}
df <- df_conditions %>%
  left_join(df_interleaf_long, 
            by = c("participant_ID", "chat_number"))
df
```
```{r}
# Keep only participants that have values
df <- df %>%
  group_by(participant_ID) %>%
  filter(!any(is.na(flow_rating))) %>%
  ungroup()

# This leaves us with the full dataset including n=42
df %>% 
  distinct(participant_ID)
```
```{r}
write.csv(df, "df_good_sessions_18_01_26.csv", row.names = FALSE)
```
